---
title: "Toronto bike share"
author: "Laurence Horton"
date: "28/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Availability
```{r packages, echo = FALSE, include = FALSE}
library(jsonlite) # For APIs
library(dplyr) # For data manipulation
library(leaflet) # For mapping
library(htmltools) # For popups on map
```

```{r get data from API, echo = FALSE, include = FALSE}
# Bike status has information about bike availability and stations (locations, etc.)
# Availability
bikestatus <-
  fromJSON("https://tor.publicbikesystem.net/ube/gbfs/v1/en/station_status")

# Stations
bikestation <-
  fromJSON("https://tor.publicbikesystem.net/ube/gbfs/v1/en/station_information")

# Convert to dataframes
status <- as_tibble(bikestatus[["data"]][["stations"]])
station <- as_tibble(bikestation[["data"]][["stations"]])
```

```{r merge, echo = FALSE, include = FALSE}
bike <- as_tibble(merge(station, status, by = "station_id"))
```

```{r create geobox, echo = FALSE, include = FALSE}
# Create geobox boundaries for stations close to me
n <- (43.66731) # north
s <- (43.66312) # south
e <- (-79.3742) # east
w <- (-79.3696) # west
```


```{r narrow search, echp = FALSE, include = FALSE}
# Keep only name of station and number of bikes available
mybike <-
  bike %>%
    rename(Station = name,
           Lat = lat,
           Lon = lon,
         Available = num_bikes_available) %>%
  select(Station, Lat, Lon, Available) %>%
  filter(Lat < n &
           Lat > s &
           Lon > e &
           Lon < w)
mybike
```

```{r drop geocordinates and rename, echo = FALSE}
availablebike <- mybike %>%
  select(-c(Lat, Lon))
availablebike
```

Map of bike availability
```{r map results, echo = FALSE}
leaflet(mybike) %>% addTiles() %>%
  addMarkers(~Lon, ~Lat, popup = ~htmlEscape(Available))
```

# Citations
## Data
Toronto Parking Authority (2021) About Bike Share Toronto, 
https://open.toronto.ca/dataset/bike-share-toronto/ published under Open Government Licence – Toronto https://open.toronto.ca/open-data-license/

## Code
Horton, L. (2021). Ontario COVID-19 Vaccinations. GitHub repository, https://github.com/laurencehorton/torontobikeshare
published under Creative Commons Zero v1.0 Universal https://github.com/laurencehorton/torontobikeshare/blob/master/LICENSE

## Packages
Cheng, J., Karambelkar, B., Xie, Y. (2019). *leaflet: Create Interactive Web Maps with the JavaScript 'Leaflet' Library*. R package version 2.0.3. https://CRAN.R-project.org/package=leaflet

Cheng, J., Sievert, C., Chang, W., Xie, Y., Allen, J. (2021). *htmltools: Tools for HTML*. R package version 0.5.1.1. https://CRAN.R-project.org/package=htmltools

Ooms, J. (2014). *The jsonlite Package: A Practical and Consistent Mapping Between JSON Data and R Objects*. arXiv:1403.2805 [stat.CO] URL https://arxiv.org/abs/1403.2805.

R Core Team (2018). *R: A language and environment for statistical computing. R Foundation for Statistical Computing*, Vienna, Austria. URL https://www.R-project.org/.

Wickham, H., François, R., Henry, L., Müller K. (2021). *dplyr: A Grammar of Data Manipulation. R package version 1.0.6*. https://CRAN.R-project.org/package=dplyr

```{r citations, echo = FALSE, include = FALSE}
citation()
citation("dplyr")
citation("jsonlite")
citation("leaflet")
citation("htmltools")
sessionInfo()
```